---
title: "Spatial Analysis Exam"
author: "Victoria Engberg Lowe, Natasha Becker Bertelsen, Vlada Caraman"
date: "06/06/2024
output:
  html_document: default
  pdf_document: default
date: "2024-05-24"
---

# Installing packages
```{r}
pacman::p_load(dplyr,
               raster,
               sf,
               readxl,
               openxlsx,
               writexl,
               mapview,
               leaflet,
               tmap,
               spdep,
               sfdep,
               lmerTest,
               tidyr,
               ggplot2,
               ggpubr,
               gridExtra,
               RColorBrewer,
               readr,
              ggspatial)
              
```


# Loading municipality data
```{r}
# set working directory
setwd('.')
getwd()

# load the municipalities data 
munic <- readRDS("Data/gadm36_DNK_2_sp.RDS")
munic_df <- st_transform(st_as_sf(munic), crs = 25832)
munic_df
```

# Cleaning the municipality data
```{r}
# getting the structure correct
munic_df <- munic_df %>% 
  dplyr::select(NAME_2, geometry) %>% 
  dplyr::rename(Municipality = NAME_2)

row.names(munic_df) <- NULL

unique(munic_df$Municipality)

# fixing the municipality names in the municipality dataframe
munic_df$Municipality[31] <- "Aarhus"
munic_df$Municipality[21] <- "Høje-Taastrup"
munic_df$Municipality[60] <- "Vesthimmerlands"

# removing municipalities where data is missing (Glostrup, Herlev, Vallensbæk) or have been excluded (Christiansø)
munic_df <- munic_df %>% 
  filter(Municipality != "Glostrup") %>% 
  filter(Municipality != "Herlev") %>% 
  filter(Municipality != "Vallensbæk") %>% 
  filter(Municipality != "Christiansø")
```

# AGE AND MUNICIPALITIES
```{r}
# load the age group data set
# load age data
age_df <- read_excel("Data/age_distribution_municipalities.xlsx")

## remove the first two rows and last two rows 
age_df <- age_df[-c(1, 2, nrow(age_df)-1, nrow(age_df)), ] 

# change column names
colnames(age_df) <- c("Age_group", "Municipality", "2023_count")

# transform variables
age_df$`2023_count` <- as.numeric(age_df$`2023_count`)

# add age group to all rows
age_df[1:99, "Age_group"] <- "20-29 years"
age_df[100:199, "Age_group"] <- "30-39 years"
age_df[199:297, "Age_group"] <- "40-49 years"
age_df[298:396, "Age_group"] <- "50-59 years"
age_df[397:495, "Age_group"] <- "60-69 years"
age_df[496:594, "Age_group"] <- "70-79 years"
age_df[595:693, "Age_group"] <- "80-89 years"

# remove younger age groups
age_df <- age_df %>% 
  filter(Age_group != "20-29 years" & Age_group != "30-39 years")

# remove municipalities where data is missing
age_df <- age_df %>% 
  filter(Municipality != "Glostrup") %>% 
  filter(Municipality != "Herlev") %>% 
  filter(Municipality != "Vallensbæk") %>% 
  filter(Municipality != "Christiansø") 

age_df <- age_df %>% 
  mutate(Municipality = case_when(
    Municipality == "Copenhagen" ~ "København",
    TRUE ~ as.character(Municipality)))

# population per municipality data set
population_munic_df <- read_excel("Data/population_per_municipality.xlsx")

# remove the first two rows and last two row 
population_munic_df <- population_munic_df[-c(1, 2, nrow(population_munic_df)-1, nrow(population_munic_df)), ] 

# change the column names
colnames(population_munic_df) <- c("Municipality", "2023_total")

# transform variables
population_munic_df$`2023_total` <- as.numeric(population_munic_df$`2023_total`) 

# removing municipalities where data is missing
population_munic_df <- population_munic_df %>% 
  filter(Municipality != "Glostrup") %>% 
  filter(Municipality != "Herlev") %>% 
  filter(Municipality != "Vallensbæk") %>% 
  filter(Municipality != "Christiansø") 

# merge all three data sets
temp_df <- merge(population_munic_df, age_df, by = "Municipality")
# add the spatial layer
age_df <- munic_df %>% 
  full_join(temp_df, join_by("Municipality" == "Municipality"))

# calculate percentage for each age group
age_df <- age_df %>% 
  mutate("2023_%" = round((`2023_count`/`2023_total`)*100,0))
```

# Plotting the age groups
```{r}
tm_shape(age_df) +
  tm_facets(by = "Age_group",
            ncol = 3,
            nrow = 3) +
  tm_fill(col = "2023_%",
          style = "pretty",
          palette = "YlGn",
          title = "% of population",
          title.size = 0.5) + 
  tm_borders(col = "black", 
             lwd = 1) +
  tm_compass(position = c("LEFT", "BOTTOM"),
             type = "rose", 
             size = 0.6) +
  tm_layout(main.title = "Age Distribution per Municipality (2023)",
            title.size = 0.3,
            bg.color = "white",
            inner.margins = c(0.1, 0.1, 0.1, 0.1),
            asp = 1.2)

```

# AGE ADJUSTED PD
## Loading in the necessary data
```{r}
# the age distributions in Denmark for the relevant age groups
age_distributions_dk <- read_xlsx("Data/age_group_national_level.xlsx")

# Parkinson data with age of diagnosis
parkinson_age <- read_excel("Data/parkinson_data_with_age_of_diagnosis.xlsx")

# total population in the municipalities
munic_pop <-read_excel("Data/people_in_the_municipality_2010_to_2023.xlsx")

# how many people in different age groups live in the municipalities
munic_age_pop <- read_excel("Data/age_distribution_absolute.xlsx")
```
## Preprocessing the data
### Age distributions
```{r}
# getting the correct age intervals
age_distributions_dk <- age_distributions_dk %>% 
  dplyr::mutate(Age_groups = case_when(
    Age_groups == "40-44 år" ~ "40-49 years",
    Age_groups == "45-49 år" ~ "40-49 years",
    Age_groups == "50-54 år" ~ "50-59 years",
    Age_groups == "55-59 år" ~ "50-59 years",
    Age_groups == "60-64 år" ~ "60-69 years",
    Age_groups == "65-69 år" ~ "60-69 years",
    Age_groups == "70-74 år" ~ "70-79 years",
    Age_groups == "75-79 år" ~ "70-79 years",
    Age_groups == "80-84 år" ~ "80-89 years", 
    Age_groups == "85 år og derover" ~ "80-89 years",
    TRUE ~ as.character(Age_groups)
  ))

# summarizing
age_distributions_dk <- age_distributions_dk %>%
  group_by(Age_groups) %>%
  mutate(Population_size_2023 = sum(Population_size_2023)) %>%
  mutate(Proportion = Population_size_2023 / Total_population) %>% 
  distinct()

# renaming the age column
age_distributions_dk <- age_distributions_dk %>% 
  rename(Age = Age_groups,
         Avg_proportion = Proportion,
         DK_population = Total_population)
```

### Parkinson with age of diagnosis
```{r}
# creating a dataset with only PD instances from 2023
PD_age_2023 <- parkinson_age %>% 
  dplyr::rename(
    Year = År,
    Municipality = Bopælskommune,
    Age = Aldersgruppe,
    PD_prevalence = `Antal patienter`) %>% 
  dplyr::filter(Year == 2023) %>% 
  dplyr::select(Municipality, Age, PD_prevalence)

# turning <5 in PD_prevalence into 2s
PD_age_2023 <- PD_age_2023 %>% 
  dplyr::mutate(PD_prevalence = dplyr::case_when(
    PD_prevalence == "<5" ~ 2,
    TRUE ~ as.numeric(PD_prevalence)))

# removing data on those aged 40 or less
PD_age_2023 <- PD_age_2023 %>% 
  dplyr::filter(!Age %in% c("15-19 år", "35-39 år"))

# changing the age group to 40-49, 50-59 etc.
PD_age_2023 <- PD_age_2023 %>% 
  dplyr::mutate(Age = case_when(
    Age == "40-44 år" ~ "40-49 years",
    Age == "45-49 år" ~ "40-49 years",
    Age == "50-54 år" ~ "50-59 years",
    Age == "55-59 år" ~ "50-59 years",
    Age == "60-64 år" ~ "60-69 years",
    Age == "65-69 år" ~ "60-69 years",
    Age == "70-74 år" ~ "70-79 years",
    Age == "75-79 år" ~ "70-79 years",
    Age == "80-84 år" ~ "80-89 years", 
    Age == "85+ år" ~ "80-89 years",
    TRUE ~ as.character(Age)
  ))

# removing municipalities where data is missing
PD_age_2023 <- PD_age_2023 %>% 
  filter(Municipality != "Glostrup") %>% 
  filter(Municipality != "Herlev") %>% 
  filter(Municipality != "Vallensbæk") %>% 
  filter(Municipality != "Christiansø")

# grouping and summarizing to get only one row per age group
PD_age_2023_summarized <- PD_age_2023 %>%
  group_by(Municipality, Age) %>%
  summarize(PD_prevalence = sum(PD_prevalence), .groups = 'drop')

# make sure all age groups are in here
PD_age_2023_summarized$Age <- factor(PD_age_2023_summarized$Age, levels = c("40-49 years", "50-59 years", "60-69 years", "70-79 years", "80-89 years"))

# create a dataframe with all combinations of Municipality and Age
all_combinations <- expand.grid(
  Municipality = unique(PD_age_2023_summarized$Municipality),
  Age = levels(PD_age_2023_summarized$Age)
)

# merge with the original dataframe, filling missing values with 0
PD_age_2023_summarized <- left_join(all_combinations, PD_age_2023_summarized, by = c("Municipality", "Age"))

# replace NA values in PD_prevalence with 0 since it means no one within that specific group was diagnosed with PD
PD_age_2023_summarized$PD_prevalence[is.na(PD_age_2023_summarized$PD_prevalence)] <- 0
```

### Population in the municipalities
```{r}
# getting the correct column names:
## extract the second row of the dataframe
new_column_names <- munic_pop[2, ]

## convert the extracted row to a character vector
new_column_names <- as.character(new_column_names)

## set the column names of the data frame to be the extracted row
colnames(munic_pop) <- new_column_names

## remove the first and last two rows from the dataframe
munic_pop <- munic_pop[-c(1:2), ]
munic_pop <- head(munic_pop, n = nrow(munic_pop) - 2)

## changing the name of the first column as it is unnamed right now
colnames(munic_pop)[1] <- "Municipality"

# making a data frame with only the data from 2023
munic_pop_2023 <- munic_pop %>% 
  dplyr::select(Municipality, "2023") %>% 
  rename(Population_2023 = "2023")

# making the population column numeric and changing <5 to 2
munic_pop_2023 <- munic_pop_2023 %>% 
  dplyr::mutate(Population_2023 = dplyr::case_when(
    Population_2023 == "<5" ~ 2,
    TRUE ~ as.numeric(Population_2023)))

# removing municipalities where data is missing
munic_pop_2023 <- munic_pop_2023 %>% 
  filter(Municipality != "Glostrup") %>% 
  filter(Municipality != "Herlev") %>% 
  filter(Municipality != "Vallensbæk") %>% 
  filter(Municipality != "Christiansø")
```


### Age distributions/populations in the municipalities
```{r}
# getting a data frame for only 2023
munic_age_pop <- munic_age_pop %>% 
  dplyr::rename(
    Municipality = municipality,
    Age = age_group,
    Population = "2023_count") %>% 
  dplyr::select(Municipality, Age, Population)
         
# renaming Copenhagen to "København"
munic_age_pop <- munic_age_pop %>% 
  mutate(Municipality = case_when(
    Municipality == "Copenhagen" ~ "København",
    TRUE ~ as.character(Municipality)))

# removing municipalities where data is missing
munic_age_pop <- munic_age_pop %>% 
  filter(Municipality != "Glostrup") %>% 
  filter(Municipality != "Herlev") %>% 
  filter(Municipality != "Vallensbæk") %>% 
  filter(Municipality != "Christiansø")
```

## Creating the relevant dataframe
```{r}
# merging pd_age_2023_summarized and munic_age_pop and munic_pop_2023 to get a dataframe with both PD prevalence and population data
PD_and_munic_df <- merge(PD_age_2023_summarized, munic_age_pop, by = c("Municipality", "Age"))
PD_and_munic_df <- merge(PD_and_munic_df, munic_pop_2023, by = "Municipality")

# renaming relevant columns
PD_and_munic_df <- PD_and_munic_df %>% 
  rename(
    Total_population = Population_2023)

# more merging to add the average age distributions for Denmark
age_adjusted_PD_df <- merge(PD_and_munic_df, age_distributions_dk, by = "Age", all.x = TRUE)

# selecting the variables we need
age_adjusted_PD_df <- age_adjusted_PD_df %>% 
  dplyr::select(Municipality, Age, Population, Avg_proportion, PD_prevalence) %>% 
  rename(Munic_age_population = Population)
```

## Calculating the age-adjusted PD prevalence
```{r}
age_adjusted_PD_df <- age_adjusted_PD_df %>% 
  mutate(Prevalence_10000 = (PD_prevalence/Munic_age_population)*10000,
         Prevalence_age_specific = Avg_proportion*Prevalence_10000) %>% 
  group_by(Municipality) %>% 
  summarize(Prevalence_standardized = sum(Prevalence_age_specific))
```

## Combining the data frame with the geometry so it can be plotted and used for hotspot analysis
```{r}
age_adjusted_PD_df <- munic_df %>% 
  full_join(age_adjusted_PD_df, join_by("Municipality" == "Municipality"))

```

## Plotting age-adjusted PD:
```{r}
# plotting
tm_shape(age_adjusted_PD_df) +
  tm_fill(col = "Prevalence_standardized",
          style = "pretty",
          palette = "YlGn",
          title = "Diagnoses per 10.000",
          title.size = 0.5) + 
  tm_borders(col = "black", 
             lwd = 1) +
  tm_scale_bar(breaks = c(0, 50, 100, 150), 
               position = c("LEFT", "TOP"),
               text.size = 0.7) +
  tm_compass(position = c("LEFT", "BOTTOM"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "N. Bertelsen, V. Caraman & V. Lowe, 2024")+
  tm_layout(main.title = "Age-Adjusted PD Prevalence (2023)",
            title.size = 0.3,
            bg.color = "white",
            inner.margins = c(0.06, 0.08, 0.08, 0.06))
```

# PESITICE USAGE
## Loading in the necessary data

```{r}
# pesticide data
pesticides_2010 <- read_xlsx(path = "Data/pesticides_data.xlsx")

# custom city to municipality mapping
city_to_munic <- read_xlsx("Data/citynames_municipalities.xlsx")
```
## Preprocessing the data
### Cleaning the pesticide data
```{r}
pesticides_2010 <- pesticides_2010 %>% 
  mutate(DosageUnit = dplyr::case_when(
    DosageUnit == 1 ~ "gram",
    DosageUnit == 2 ~ "kilo",
    DosageUnit == 3 ~ "mililiters",
    DosageUnit == 4 ~ "liters",
    DosageUnit == 5 ~ "tablets",
    TRUE ~ as.character(DosageUnit))) %>% 
  mutate(DosageQuantity = dplyr::case_when(
    NoPesticides == "TRUE" ~ 0,
    TRUE ~ as.numeric(DosageQuantity)))
```

### Cleaning the city to municipality data
```{r}
city_to_munic <- city_to_munic %>% 
  dplyr::select(City, Municipality)
```

### Making a mapping between city municipality to use on the pesticde dataframe 
```{r}
# Create a vector mapping city names to municipality names
city_to_municipality <- setNames(city_to_munic$Municipality, city_to_munic$City)

# Using it to make a new column named "Municipality" in pesticides_2010
pesticides_2010 <- pesticides_2010 %>%
  mutate(Municipality = city_to_municipality[City])

pesticides_2010 <- pesticides_2010 %>% 
  dplyr::select(City, Municipality, NoPesticides, DosageQuantity, DosageUnit)
```

## Getting dataframe for the pesticide data ready for plotting and further analysis
```{r}
# remove data on the tablets as they are insignificant
pesticides_df <- pesticides_2010 %>% 
  dplyr::filter(is.na(DosageUnit) | DosageUnit == "kilo" | DosageUnit == "liters")

unique(pesticides_df$DosageUnit)

# making a total use where liters and kiloas are combined and 1 liter = 1 kg is assumed per the instructions from Miljøstyrelsen
pesticides_df <- pesticides_df %>% 
  group_by(Municipality) %>%
  dplyr::summarize(TotalUse = round(sum(DosageQuantity), digits = 2))

# Removing municipalities where data is missing and "nordfyn" as that is not present in the other dataframes
pesticides_df <- pesticides_df %>% 
  dplyr::filter(Municipality != "Glostrup") %>% 
  dplyr::filter(Municipality != "Herlev") %>% 
  dplyr::filter(Municipality != "Vallensbæk") %>% 
  dplyr::filter(Municipality != "Christiansø") %>% 
  dplyr::filter(Municipality != "Nordfyn")

pesticides_df <- munic_df %>% 
  dplyr::full_join(pesticides_df, dplyr::join_by("Municipality" == "Municipality"))
```


## Plotting pesticide usage from pesticide_df
```{r}
tm_shape(pesticides_df) +
  tm_fill(col = "TotalUse", 
          style = "pretty",
          palette = "YlGn",  
          title = "Kg used",
          title.size = 0.2) +  
  tm_borders(col = "black", 
             lwd = 1) +
  tm_scale_bar(breaks = c(0, 50, 100, 150), 
               position = c("LEFT", "TOP"),
               text.size = 0.7) +
  tm_compass(position = c("LEFT", "BOTTOM"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "N. Bertelsen, V. Caraman & V. Lowe, 2024") +
  tm_layout(main.title = "Pesticide Usage in Kilos (2010-2011)",
            title.size = 0.1,
            bg.color = "white",
            inner.margins = c(0.15, 0.15, 0.15, 0.15),
            legend.position = c("RIGHT", "TOP"))
```

# CROP DENSITY
## Importing the data

```{r}
# read in the farming per municipality data
farming_df <- read_xlsx("Data/farming_municipalities_km2.xlsx")

# read in the total area per municipality data
munic_area_df <- read_xlsx("Data/municipality_area_2011.xlsx")
```
## Preprocessing the data 
### Farming area

```{r}
# Farming data set
# remove the first two columns
farming_df <- farming_df[, -c(1, 2)]

# remove the first three rows
farming_df <- farming_df[-c(1:3), ]

# change column names
colnames(farming_df) <- c("Municipality", "2011")

# transform variable
farming_df$`2011` <- as.numeric(farming_df$`2011`) 

# rename columns
farming_df <- farming_df %>% 
  rename(farming_area_in_km2_2011 = "2011")

# remove municipalities where data is missing
farming_df <- farming_df %>% 
  filter(Municipality != "Glostrup") %>% 
  filter(Municipality != "Herlev") %>% 
  filter(Municipality != "Vallensbæk") %>% 
  filter(Municipality != "Christiansø")
```

### The total area
```{r}
# clean munic_area_df prior to merging
munic_area_df$`Area in km2 in 2011` <- as.numeric(munic_area_df$`Area in km2 in 2011`)

munic_area_df <- munic_area_df %>% 
  rename(Area_in_km2_2011 = `Area in km2 in 2011`)

# remove municipalities where data is missing
munic_area_df <- munic_area_df %>% 
  filter(Municipality != "Glostrup") %>% 
  filter(Municipality != "Herlev") %>% 
  filter(Municipality != "Vallensbæk") %>% 
  filter(Municipality != "Christiansø")
```

### Merging the two dataframes
```{r}
# merging the two dataframes
crop_density_df <- merge(farming_df, munic_area_df, by = "Municipality")

# adding the spatial layer
crop_density_df <- munic_df %>% 
  full_join(crop_density_df, join_by("Municipality" == "Municipality"))
```

## Calculating crop density
```{r}
crop_density_df <- crop_density_df %>% 
  mutate(crop_density = farming_area_in_km2_2011/Area_in_km2_2011) %>% 
  dplyr::select(Municipality, Area_in_km2_2011, farming_area_in_km2_2011, crop_density, geometry)
```

## Plotting crop density
```{r}
tm_shape(crop_density_df) +
  tm_fill(col = "crop_density", 
          style = "pretty",
          palette = "YlGn",  
          title = "Crop density",
          title.size = 0.2) +  
  tm_borders(col = "black", 
             lwd = 1) +
  tm_scale_bar(breaks = c(0, 50, 100, 150), 
               position = c("LEFT", "TOP"),
               text.size = 0.7) +
  tm_compass(position = c("LEFT", "BOTTOM"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "N. Bertelsen, V. Caraman & V. Lowe, 2024") +
  tm_layout(main.title = "Crop Density (2011)",
            title.size = 0.1,
            bg.color = "white",
            inner.margins = c(0.15, 0.15, 0.15, 0.15),
            legend.position = c("RIGHT", "TOP"))
```

# Checking if the geometries match
```{r}
identical(st_geometry(munic_df), st_geometry(crop_density_df))
identical(st_geometry(munic_df), st_geometry(pesticides_df))
identical(st_geometry(munic_df), st_geometry(age_adjusted_PD_df))
```
# SPATIAL AUTOCORRELATIONS AND HOT SPOT ANALYSIS
## Defining neighbors for each municipality

```{r}
# set seed
set.seed(123)
# get center points of each municipality
mun_centers <- st_centroid(munic_df$geometry, of_largest_polygon = TRUE)

# make neighbors list from k neighbors
nb_k2 <- knn2nb(knearneigh(mun_centers, k = 3))

# convert the k-nearest neighbors list to a weights list
w_knn_k2 <- nb2listw(nb_k2, style = "B")

```

# Age-adjusted PD
```{r}
globalG.test(age_adjusted_PD_df$Prevalence_standardized, w_knn_k2)
age_adjusted_PD_df_knn <- age_adjusted_PD_df %>%
  mutate(
    age_adjusted_PD_df_lag = lag.listw(w_knn_k2, age_adjusted_PD_df$Prevalence_standardized)
  )
```
## Local Gi test fOr age-adjusted PD
```{r}
# test for local spatial autocorrelation 
age_adjusted_PD_df_hot_spots <- age_adjusted_PD_df_knn %>%
  mutate(
    Gi = local_g_perm(age_adjusted_PD_df$Prevalence_standardized, nb_k2, w_knn_k2$weights, nsim = 999)
  )

# unnest Gi
age_adjusted_PD_df_hot_spots <- age_adjusted_PD_df_hot_spots %>% 
  unnest(cols = c(Gi))

```

## Plotting the age-adjusted PD
```{r}
plot_pd <- age_adjusted_PD_df_hot_spots %>%  
  dplyr::select(gi, p_folded_sim) %>% 
  mutate(
   classification = case_when(
      gi > 0 & p_folded_sim <= 0.05 ~ "Hot",
      gi > 0 & p_folded_sim <= 0.1 ~ "Somewhat hot",
      gi < 0 & p_folded_sim <= 0.05 ~ "Cold",
      gi < 0 & p_folded_sim <= 0.1 ~ "Somewhat cold",
      TRUE ~ "Insignificant"
    ),
    classification = factor(
      classification,
      levels = c("Hot", "Somewhat hot",
                 "Insignificant",
                 "Somewhat cold", "Cold")
    )
  ) %>%  
 ggplot(aes(fill = classification)) +
  geom_sf(color = "black", lwd = 0.1) +
  scale_fill_manual(values = c(
    "Hot" = "#00441b",    # Dark green
    "Somewhat hot" = "#bae4b3", # Lighter green
    "Insignificant" ="#f7fcf5",      # Almost white 
    "Somewhat cold" = "#fdae6b",  # Light orange 
    "Cold" = "#d94801"      # Dark orange
  )) +
  theme_void() +
  labs(
    fill = "Hot Spot Classification",
    title = "Hot Spot Analysis of Age-Adjusted PD (2023)",
    caption = "N. Bertelsen, V. Caraman & V. Lowe, 2024") +
    theme(plot.caption = element_text(hjust = 0.5)) +
    annotation_north_arrow(
    location = "tl",  # Bottom left
    which_north = "true",
    pad_x = unit(0.2, "in"),
    pad_y = unit(0.2, "in"),
    style = north_arrow_fancy_orienteering,
    height = unit(0.9, "cm"),  
    width = unit(0.9, "cm")  
  )

plot_pd

```

# Pesticide usage
```{r}
globalG.test(pesticides_df$TotalUse, w_knn_k2)
pesticides_df_knn <- pesticides_df %>%
  mutate(
    pesticides_lag = lag.listw(w_knn_k2, pesticides_df$TotalUse)
  )

```

## Local Gi test for pesticide usage
```{r}
# test for local spatial autocorrelation 
pesticides_hot_spots <- pesticides_df_knn %>%
  mutate(
    Gi = local_g_perm(pesticides_df$TotalUse, nb_k2, w_knn_k2$weights, nsim = 999)
  )

# unnest Gi results 
pesticides_hot_spots <- pesticides_hot_spots %>% 
  unnest(cols = c(Gi))

```

## Plotting the pesticide usage
```{r}
plot_pesticides  <- pesticides_hot_spots %>%  
  dplyr::select(gi, p_folded_sim) %>% 
  mutate(
   classification = case_when(
      gi > 0 & p_folded_sim <= 0.05 ~ "Hot",
      gi > 0 & p_folded_sim <= 0.1 ~ "Somewhat hot",
      gi < 0 & p_folded_sim <= 0.05 ~ "Cold",
      gi < 0 & p_folded_sim <= 0.1 ~ "Somewhat cold",
      TRUE ~ "Insignificant"
    ),
    classification = factor(
      classification,
      levels = c("Hot", "Somewhat hot",
                 "Insignificant",
                 "Somewhat cold", "Cold")
    )
  ) %>%  
 ggplot(aes(fill = classification)) +
  geom_sf(color = "black", lwd = 0.1) +
  scale_fill_manual(values = c(
    "Hot" = "#00441b",    # Dark green
    "Somewhat hot" = "#bae4b3", # Lighter green
    "Insignificant" ="#f7fcf5",      # Almost white 
    "Somewhat cold" = "#fdae6b",  # Light orange 
    "Cold" = "#d94801"      # Dark orange
  )) +
  theme_void() +
  labs(
    fill = "Hot Spot Classification",
    title = "Hot Spot Analysis of Pesticide Usage (2010-2011)",
    caption = "N. Bertelsen, V. Caraman & V. Lowe, 2024")+
    theme(plot.caption = element_text(hjust = 0.5)) +
    annotation_north_arrow(
    location = "tl",  # Bottom left
    which_north = "true",
    pad_x = unit(0.2, "in"),
    pad_y = unit(0.2, "in"),
    style = north_arrow_fancy_orienteering,
    height = unit(0.9, "cm"),  
    width = unit(0.9, "cm")  
  )


plot_pesticides
```

# Crop density
```{r}
globalG.test(crop_density_df$crop_density, w_knn_k2)
crop_density_knn <- crop_density_df %>%
  mutate(
    farm_lag = lag.listw(w_knn_k2, crop_density_df$crop_density)
  )

```

## Local Gi test for crop density
```{r}
# test for local spatial autocorrelation 
crop_density_hot_spots <- crop_density_knn %>%
  mutate(
    Gi = local_g_perm(crop_density_df$crop_density, nb_k2, w_knn_k2$weights, nsim = 999)
  )

# unnest Gi results
crop_density_hot_spots <- crop_density_hot_spots %>% 
  unnest(cols = c(Gi))

```

## Plotting the crop density
```{r}
crop_density_hot_spots %>%  
  dplyr::select(gi, p_folded_sim) %>% 
  mutate(
    classification = case_when(
      gi > 0 & p_folded_sim <= 0.05 ~ "Hot",
      gi > 0 & p_folded_sim <= 0.1 ~ "Somewhat hot",
      gi < 0 & p_folded_sim <= 0.05 ~ "Cold",
      gi < 0 & p_folded_sim <= 0.1 ~ "Somewhat cold",
      TRUE ~ "Insignificant"
    ),
    classification = factor(
      classification,
      levels = c("Hot", "Somewhat hot",
                 "Insignificant",
                 "Somewhat cold", "Cold")
    )
  ) %>%  
 ggplot(aes(fill = classification)) +
  geom_sf(color = "black", lwd = 0.1) +
  scale_fill_manual(values = c(
    "Hot" = "#00441b",    # Dark green
    "Somewhat hot" = "#bae4b3", # Lighter green
    "Insignificant" ="#f7fcf5",      # Almost white 
    "Somewhat cold" = "#fdae6b",  # Light orange 
    "Cold" = "#d94801"      # Dark orange
  )) +
  theme_void() +
  labs(
    fill = "Hot Spot Classification",
    title = "Hot Spot Analysis of Crop Density (2011)",
    caption = "N. Bertelsen, V. Caraman & V. Lowe, 2024") +
    theme(plot.caption = element_text(hjust = 0.5)) +
    annotation_north_arrow(
    location = "tl",  # Bottom left
    which_north = "true",
    pad_x = unit(0.2, "in"),
    pad_y = unit(0.2, "in"),
    style = north_arrow_fancy_orienteering,
    height = unit(0.9, "cm"),  
    width = unit(0.9, "cm")  
  )


```

## Creating a dataframe for the model
```{r}
pesticides_gi_df <- as.data.frame(pesticides_hot_spots)
pesticides_gi_df <- pesticides_gi_df %>% rename(gi_pesticides = gi)
crop_density_gi_df <- as.data.frame(crop_density_hot_spots)
crop_density_gi_df <- crop_density_gi_df %>% rename(gi_crop_density = gi)
age_adjusted_PD_gi_df <- as.data.frame(age_adjusted_PD_df_hot_spots)
age_adjusted_PD_gi_df <- age_adjusted_PD_gi_df %>% rename(gi_PD = gi)

# select relevant columns
pesticides_gi_df <- pesticides_gi_df  %>% 
  dplyr::select(Municipality, gi_pesticides)

crop_density_gi_df <- crop_density_gi_df %>% 
  dplyr:: select(Municipality, gi_crop_density)

age_adjusted_PD_gi_df <- age_adjusted_PD_gi_df %>% 
  dplyr:: select(Municipality, gi_PD)

# merge the data frames
df_model <- merge(pesticides_gi_df, crop_density_gi_df, by = "Municipality")
df <- merge(age_adjusted_PD_gi_df, df_model, by = "Municipality", all.x = TRUE)


```


# LINEAR MODEL
```{r}
# defining the model
model <- lm(gi_PD ~ gi_pesticides + gi_crop_density, data = df)
summary(model)


## checking residuals
# extract residuals and predicted values
df_fitted <- df %>% 
  mutate(predicted = predict(model), resid = residuals(model))

## checking for normality
# density plot
density_plot <- df_fitted %>% 
  ggplot() +
  aes(x = resid) +
  geom_density() +
  theme_pubr() +
  labs(x = "Residuals", y = "Density", title = "Density of residuals")

# QQ-plot
qq_plot <- ggqqplot(df_fitted, x = "resid") +
  labs(title = "QQ-plot of residuals")

# homoscedasticity and linearity of residuals
pred_vs_res <- df_fitted %>% 
  ggplot() +
  aes(x = predicted, y = resid) +
  geom_point() + 
  geom_hline(yintercept = 0) +
  theme_pubr() +
  labs(title = "Fitted vs. residuals", x = "Fitted values", y = "Residuals")

grid.arrange(density_plot, qq_plot, pred_vs_res, ncol = 3)

# checking for spatial autocorrelation using Moran's I
moran.test(df_fitted$resid,
           nb2listw(nb_k2, style = "W",zero.policy=TRUE),
           zero.policy=TRUE)
  
```


